---
title: "gcamInvest"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gcamInvest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-------------------------->
<!-------------------------->
# Read GCAM Data and Calculate Electricity Investment
<!-------------------------->
<!-------------------------->

```plutus::gcamInvest``` integrates the functionality of reading CGAM data and calculating stranded assets and electricity investments in power sector in one place. This function will return a list containing:

- data: a dataframe with the original data
- dataAggParam: a dataframe with the data aggregated to the parameter
- dataAggClass1: a dataframe with the data aggregated to class 1
- dataAggclass2: a dataframe with the data aggregated to class 2
- scenarios: A list of the scenarios
- queries: A list of the queries used to extract the data

The details of arguments and their default values in ```plutus::gcamInvest``` can be found in the [gcamInvest reference page](https://jgcri.github.io/plutus/reference/gcamInvest.html). The following sections provide step-by-step instructions on using ```plutus::gcamInvest```.

```{r eval=F}
# Default argument values
plutus::gcamInvest(gcamdatabase = NULL,
                   queryFile = NULL,
                   reReadData = T,
                   dataProjFile = paste(getwd(), "/outputs/dataProj.proj", sep = ""),
                   gcamdataFile = NULL,
                   scenOrigNames = 'All',
                   scenNewNames = NULL,
                   regionsSelect = NULL,
                   dirOutputs = paste(getwd(), "/outputs", sep = ""),
                   folderName = NULL,
                   nameAppend = "",
                   saveData = T)
```

## Read Project File

```plutus::gcamInvest``` is able to read .proj file from GCAM output by providing the path to the .proj file. Plutus also includes an example .proj dataset ```plutus::exampleGCAMproj```.

```{r eval=F}
library(plutus)

invest <- plutus::gcamInvest(# dataProjFile = path_to_projfile,
                             dataProjFile = plutus::exampleGCAMproj)

# Explore the list returned from plutus::gcamInvest
df <- invest$data; df
dfParam <- invest$dataAggParam; dfParam
dfClass1 <- invest$dataAggClass1; dfClass1
dfScenario <- invest$scenarios; dfScenario
dfQuery <- invest$queries; dfQuery

```

## Read GCAM Database

```plutus::gcamInvest``` can directly read GCAM database. Assign the path of GCAM database to ```gcamdatabase``` argument.


```{r eval=F}
library(plutus)

# provide path to the desired GCAM database folder.
path_to_gcamdatabase <- 'E:/gcam-core-gcam-v5.3/output/databse_basexdb'
invest <- plutus::gcamInvest(gcamdatabase = path_to_gcamdatabase)

```

<!-------------------------->
<!-------------------------->
# Read and Subset GCAM Data
<!-------------------------->
<!-------------------------->

```plutus::gcamInvest``` provides options to subset GCAM data by scenario and region.


```{r eval=F}
library(plutus)

invest <- plutus::gcamInvest(dataProjFile = plutus::exampleGCAMproj,
                             scenOrigNames = c('Reference', 'Impacts', 'Policy'),
                             scenNewNames = c('Reference', 'Climate Impacts', 'Climate Policy'),
                             regionsSelect = c('USA', 'Argentina'))

df <- invest$data; df
dfParam <- invest$dataAggParam; dfParam
dfClass1 <- invest$dataAggClass1; dfClass1

```

<!-------------------------->
<!-------------------------->
# Input Data and Assumptions
<!-------------------------->
<!-------------------------->

Plutus provides default data and assumptions from GCAM v5.3, which can be found in Your-GCAM-5.3-Folder/input/gcamdata/. All CSV files associated with different data categories and assumptions are listed in [Table 1](#table1).

<a name="table1"></a>
**Table 1:** Data and assumption files

| Data or Assumption | Technology | Region | Data File |
|---|---|---|---|
| Overnight capital costs | Electricity generation technologies | Global | L2233.GlobalIntTechCapital_elec.csv <br /> L2233.GlobalTechCapital_elecPassthru.csv |
| Overnight capital costs | Cooling technologies | Global | L2233.GlobalIntTechCapital_elec_cool.csv <br /> L2233.GlobalTechCapital_elec_cool.csv |
| Capacity factors | Electricity generation technologies | Global | L223.GlobalTechCapFac_elec.csv |
| Capacity factors | Intermittant technologies | Regional | L223.StubTechCapFactor_elec.csv |
| Lifetime and steepness| Electricity generation technologies | Global | A23.globaltech_retirement.csv |

```plutus::gcamInvest``` allows users to input their own data associated with their GCAM runs. Users can update values in any CSV file. Please keep the file name and the format of each updated file unchanged. To input your own data, specify the path to the folder that contains all CSV files listed in [Table 1](#table1) in the ```gcamdataFile``` argument.

```{r eval=F}
library(plutus)

invest <- plutus::gcamInvest(dataProjFile = plutus::exampleGCAMproj,
                             gcamdataFile = 'E:/gcam-core-gcam-v5.3/input/gcamdata')
```

<!-------------------------->
<!-------------------------->
# Output Options
<!-------------------------->
<!-------------------------->

```plutus::gcamInvest``` automatically saves outputs, such as .proj files, queries, and output data tables in default path based on the working directory. This section will introduce arguments in ```plutus::gcamInvest``` that controls different output options.

## Re-read Data

```plutus::gcamInvest``` automatically save a copy of .proj file (default name: dataProj.proj) in the output directory after extracting GCAM output data if provided a GCAM database folder using ```gcamdatabase``` argument. Reload the same (subsetted) GCAM dataset can be much faster by using the automatically save dataProj.proj file. Users can choose to turnoff auto save function by specifying argument to ```reReadData = F```.

```{r eval=F}
library(plutus)

path_to_gcamdatabase <- 'E:/gcam-core-gcam-v5.3/output/databse_basexdb'
invest <- plutus::gcamInvest(gcamdatabase = path_to_gcamdatabase,
                             reReadData = F) # Default is reReadData = T
```


## Output Directory

It is easy to change output folder names and names appended to the output data tables.

  - ```folderName``` creates a folder with the specified name under output directory 
  - ```nameAppend``` appends the specified string at the end of each output data table file 

```{r eval=F}
library(plutus)

invest <- plutus::gcamInvest(dataProjFile = plutus::exampleGCAMproj,
                             # dirOutputs = Your-desired-output-path, # Default is paste(getwd(), "/outputs", sep = "")
                             folderName = 'USA', # Default is fodlerName = NULL
                             nameAppend = '_Invest') # Default is nameAppend = ''
```

You can also choose not to save any output by setting argument to ```saveData = F```.
```{r eval=F}
library(plutus)

invest <- plutus::gcamInvest(dataProjFile = plutus::exampleGCAMproj,
                             saveData = F) # Default is saveData = T
```
